Think deep for reef resilience

### 2025-08-12 paper/paper_draft.qmd
Fix ECS units and sources in table

Projections using Shared Socio-economic Pathway (SSP) 2-4.5 was used to represent the
plausible emission pathway. It represents a moderate emissions scenario with continued CO2
emissions comparable to present day until 2050, after which emissions gradually decline
[@hausfather2020; IPCC -@climate2023]. The selected GCMs represent 2.5°C to 4.31°C **\[TI:
Use K or Celcius?\]** of global atmospheric warming in the 2081-2100 period relative to
1850-1900. The use of this range of GCMs therefore allow for a range of possible climate
futures to be explored, capturing the uncertainty in future responses to CO2 increases
(@sfig-ecs; @fig-methods-dhws; [@grose2023; @shepherd2018]) and enables the assessment of
model projected coral reef behaviour over a broad range of independently derived climate
futures [@abramowitz2019].

@@

::: {#tbl-gcms}
| Global Climate Model source | ECS value (K) | Interpreted relative severity | Results found in |
|-----------------------|-----------------------|------------------------|-----------------------|
| EC-Earth3-Veg | 4.30 | High | **insert repo citation** |
| ACCESS-CM2 | 4.66 | High | **insert repo citation** |
| GFDL-CM4 | 4.1 | High | **insert repo citation** |
| ACCESS-ESM1-5 | 3.87 | Intermediate | Main text |
| NorESM2-MM | 2.5 | Low | **insert repo citation** |

: Information for each GCM used in this investigation including Equilibrium Climate
Sensitivity (ECS) value, interpreted relative severity and where results are individual GCM
results are found. GCMs in context with assessed plausible ranges of ECS values are
displayed in [@sfig-ecs]. ACCESS-ESM1-5 has an ECS closest to the midpoint of the
very-likely ECS range and was taken as an intermediate model with results displayed in the
main text. ECS values obtained from: [@sentman2025; @grose2023; @seland2020; @ziehn2020;
@wyser2020]. **\[TI: Ben, I would add this to the table directly in the ECS value column
(ECS value and source). The table should also be ordered by ECS value (high to low)\]**
:::

@@

NorESM2-MM has the lowest Equilibrium Climate Sensitivity (2.5°C) and reefs exhibit
continuing recovery and decline dynamics up to 2075 under this possible climate future.
Recovery phases are predominantly seen in the ‘high’ cover cluster. The ‘medium’ cluster,
and to a lesser degree the ‘low’ cluster, also exhibit periods of recovery and increasing
proportions of coral cover up to 2045. Under NorESM2-MM the Far Northern management area has
lower CID distances between low and high cover clusters than in other management areas
[@tbl-cluster-cid], this is reflected in the low degree of cover maintainence
[@fig-gcm-cover-trajectories]. ACCESS-ESM1-5 has an intermediate ECS value of 3.87°C which
results in intermediate levels of coral cover maintenance (between those of lower and higher
ECS GCMs) [@fig-gcm-cover-trajectories], reflected in intermediate CID distance values
within most management areas [@tbl-cluster-cid]. In Townsville-Whitsunday and Far Northern
management areas, reefs in the high cluster are able to maintain coral cover levels around
20% to 2050 under ACCESS-ESM1-5. Under ACCESS-CM2 (4.66K), EC-Earth3-Veg (4.30K) and
GFDL-CM4 (4.1K) GCMs cover trajectories are largely declining in all management areas,
reflected in lower CID distance values than other GCMs [@tbl-cluster-cid].

>>>>

Projections using Shared Socio-economic Pathway (SSP) 2-4.5 was used to represent the
plausible emission pathway. It represents a moderate emissions scenario with continued CO2
emissions comparable to present day until 2050, after which emissions gradually decline
[@hausfather2020; IPCC -@climate2023]. The selected GCMs represent 2.5°C to 4.31°C of global atmospheric warming in the 2081-2100 period relative to
1850-1900. The use of this range of GCMs therefore allow for a range of possible climate
futures to be explored, capturing the uncertainty in future responses to CO2 increases
(@sfig-ecs; @fig-methods-dhws; [@grose2023; @shepherd2018]) and enables the assessment of
model projected coral reef behaviour over a broad range of independently derived climate
futures [@abramowitz2019].

@@

::: {#tbl-gcms}
| Global Climate Model source | ECS value (°C) | Interpreted relative severity | Results found in |
|-----------------------|-----------------------|------------------------|-----------------------|
| ACCESS-CM2 | 4.66 [@grose2023]| High | **insert repo citation** |
| EC-Earth3-Veg | 4.30 [@wyser2020]| High | **insert repo citation** |
| GFDL-CM4 | 4.1 [@sentman2025]| High | **insert repo citation** |
| ACCESS-ESM1-5 | 3.87 [@ziehn2020]| Intermediate | Main text |
| NorESM2-MM | 2.5 [@seland2020]| Low | **insert repo citation** |

: Information for each GCM used in this investigation including Equilibrium Climate
Sensitivity (ECS) value, interpreted relative severity and where results are individual GCM
results are found. GCMs in context with assessed plausible ranges of ECS values are
displayed in [@sfig-ecs]. ACCESS-ESM1-5 has an ECS closest to the midpoint of the
very-likely ECS range and was taken as an intermediate model with results displayed in the
main text.
:::

@@

NorESM2-MM has the lowest Equilibrium Climate Sensitivity (2.5°C) and reefs exhibit
continuing recovery and decline dynamics up to 2075 under this possible climate future.
Recovery phases are predominantly seen in the ‘high’ cover cluster. The ‘medium’ cluster,
and to a lesser degree the ‘low’ cluster, also exhibit periods of recovery and increasing
proportions of coral cover up to 2045. Under NorESM2-MM the Far Northern management area has
lower CID distances between low and high cover clusters than in other management areas
[@tbl-cluster-cid], this is reflected in the low degree of cover maintainence
[@fig-gcm-cover-trajectories]. ACCESS-ESM1-5 has an intermediate ECS value of 3.87°C which
results in intermediate levels of coral cover maintenance (between those of lower and higher
ECS GCMs) [@fig-gcm-cover-trajectories], reflected in intermediate CID distance values
within most management areas [@tbl-cluster-cid]. In Townsville-Whitsunday and Far Northern
management areas, reefs in the high cluster are able to maintain coral cover levels around
20% to 2050 under ACCESS-ESM1-5. Under ACCESS-CM2 (4.66°C), EC-Earth3-Veg (4.30°C) and
GFDL-CM4 (4.1°C) GCMs cover trajectories are largely declining in all management areas,
reflected in lower CID distance values than other GCMs [@tbl-cluster-cid].
